MCTS Applied to Co-operative Problems

### 2015-07-03 Report/ceec.tex

The problems are designed to require Agents to co-operate in order to achieve their goals. Unlike other environments where \gls{mcts} can charge off achieving its own goals, this domain requires the agents to sometimes sit patiently on a button for another Agent to do something else.

@@

With this observation made - the computational budget was increased ten-fold, and the \gls{mcts} agents performed significantly better. Typical behaviour would be near optimal with significantly less random sections of movement. This can likely be attributed to the increased rollout and uct border depths allowing more foresight on such a small puzzle.
\subsection{Tournament}
For the main data collection, we ran a round robin tournament of maps and Agents.

@@

\subsubsection{\gls{ai} Controllers}
A small set of \gls{ai} Controllers were created in order to attempt to solve the problem domain. None of the agents have the ability to communicate with another agent.
\paragraph{Random}
The random controller simply chooses one of the possible 5 actions. This is one of the simplest to implement and run.
\paragraph{Keyboard}
The keyboard controller takes input from the keyboard and translates that into a single Action. Once an action is polled, it is reset to the No Op action unless the keyboard is used again. 
\paragraph{\gls{mcts}}
The \gls{mcts} controller is a simple implementation of \gls{uct} - with a fixed iteration, \gls{uct} Border and rollout border. No knowledge about the game is provided and the assumption is made that the other player will play randomly. The fixed rollout border makes a significant improvement in iterations per decision (On the machine from 1-3 to 500-600 in 40ms). This greatly improves the ability of \gls{mcts} to make informed decisions. The score at the end of a rollout is taken from the game state - so if \gls{mcts} doesn't see any player reach a goal, all branches will be equal.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{level1vsleve1e}
\caption{Graph showing the average score obtained by each Agent across level 1(Blue) and level 1E (Orange)}
\label{1v1eS}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{level1vsleve1eTicks}
\caption{Graph showing the average ticks taken for each AI Agent across level 1(Blue) and level 1E (Orange)}
\label{1v1eT}
\end{figure}

Figure \ref{1v1eS} shows the effect that having the complex \emph{Buttons} and \emph{Doors} in a level has on the score an Agent achieves. Figure \ref{1v1eT} demonstrates the vast improvement in pathfinding ability to a goal in the VariGA compared to the GAController.

@@

\pagebreak
\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{level5}
\caption{The initial starting state for level5.txt}
\label{Level5State}
\end{figure}
Level 5 - shown in Figure \ref{Level5State} - poses a particular problem for many of the agents (see Figure \ref{avgScoreMap5}). The asymmetric nature of the level, and the delayed reward caused great difficulty for the Agents. Relying on two button presses and the other agent to get through the open doors - in order - greatly reduced performance. The top agent, High MCTS, only managed an average score of $0.44$ compared to its total average of just over 0.8.

\begin{figure}[ht]
\centering
\includegraphics[trim=0cm 0cm 5cm 3cm, clip=true, width = \linewidth]{avgScoreOverMap5}
\caption{Average Score for each AI agent over Map 5}
\label{avgScoreMap5}
\end{figure}

\section{Conclusions}
In this paper, we found that a strong MCTS player can solve simple cooperative problems without requiring communication between agents.
\bibliographystyle{IEEEtran}
\bibliography{ceec}

>>>>

The problems are designed to require \emph{Agents} to co-operate in order to achieve their goals. Unlike other environments where \gls{mcts} can charge off achieving its own goals, this domain requires the agents to sometimes sit patiently on a button for another Agent to do something else.

@@

With this observation made - the computational budget was increased ten-fold, and the \gls{mcts} agents performed significantly better. Typical behaviour would be near optimal with significantly less random sections of movement. This can likely be attributed to the increased rollout and \gls{uct} border depths allowing more foresight on such a small puzzle.
\subsection{Tournament}
For the main data collection, we ran a round robin tournament of maps and Agents.

@@

\subsubsection{\gls{ai} Controllers}
A small set of \gls{ai} Controllers were created in order to attempt to solve the problem domain. None of the agents have the ability to communicate with another agent.
\paragraph{Random}
The random controller simply chooses one of the possible 5 actions. This is one of the simplest to implement and run.
\paragraph{Keyboard}
The keyboard controller takes input from the keyboard and translates that into a single Action. Once an action is polled, it is reset to the No Op action unless the keyboard is used again. 
\paragraph{\gls{mcts}}
The \gls{mcts} controller is a simple implementation of \gls{uct} - with a fixed iteration, \gls{uct} Border and rollout border. No knowledge about the game is provided and the assumption is made that the other player will play randomly. The fixed rollout border makes a significant improvement in iterations per decision (from 1-3 to 500-600 in 40ms \footnote{Intel Core i5-3570, 8GB RAM, Windows 7 Enterprise 64bit}). This greatly improves the ability of \gls{mcts} to make informed decisions. The score at the end of a rollout is taken from the game state - so if \gls{mcts} doesn't see any player reach a goal, all branches will be equal.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{level1vsleve1e}
\caption{Graph showing the average score obtained by each \emph{Agent} across level 1(Blue) and level 1E (Orange)}
\label{1v1eS}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{level1vsleve1eTicks}
\caption{Graph showing the average ticks taken for each \gls{ai} \emph{Agent} across level 1(Blue) and level 1E (Orange)}
\label{1v1eT}
\end{figure}

Figure \ref{1v1eS} shows the effect that having the complex \emph{Buttons} and \emph{Doors} in a level has on the score an Agent achieves. Figure \ref{1v1eT} demonstrates the vast improvement in path finding ability to a goal in the VariGA compared to the GAController. This is expected due to the ability to make individual actions as compared to the GAController which is only able to make 3 identical turns at a time leading to convoluted paths compared to optimum.

@@

\pagebreak
\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{level5}
\caption{The initial starting state for level5.txt}
\label{Level5State}
\end{figure}
Level 5 - shown in Figure \ref{Level5State} - poses a particular problem for many of the agents (see Figure \ref{avgScoreMap5}). The asymmetric nature of the level, and the delayed reward caused great difficulty for the Agents. Relying on two button presses and the other agent to get through the open doors - in order - greatly reduced performance. The top agent, High MCTS, only managed an average score of $0.44$ compared to its total average of just over $0.8$.

\begin{figure}[ht]
\centering
\includegraphics[ width = \linewidth]{avgScoreOverMap5}
\caption{Average Score for each AI agent over Map 5}
\label{avgScoreMap5}
\end{figure}

\section{Conclusions}
In this paper, we found that a strong MCTS player can solve simple cooperative problems without requiring communication between agents. We also found that a number of other \gls{ai} techniques experienced difficulty when the problem became cooperative. Further work is required in this area.

\subsection{Further Work}
There is scope for more work in a number of areas. More work can be conducted in the writing of AI's to use communication in order to complete these challenges better amongst themselves. A framework for communication would be devised - with the requirement for fairly \gls{ggp} restrictions in mind.

More work can also be done on experimenting with the problem domain itself. Expanding the object types to include different techniques such as pickup-able objects that can be placed on \emph{Buttons}, \emph{Buttons} that remain active for a period of time after vacating them as well as potentially \emph{Doors} that will open permanently if all it's \emph{Buttons} are activated. The scope here is fairly limitless.

\bibliographystyle{IEEEtran}
\bibliography{ceec}
